`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys

### 2019-07-03 analysis/paper/sim_survey_paper.Rmd
Split the case study table into tables on the core functions

The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, as well as functions for simulating surveys with a range of sampling stratigies and intensities (Table \@ref(tab:fun)). The equations behind these functions are detailed in the sections below. **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. Package documentation has been published online using **`pkgdown`** (https://paulregular.github.io/SimSurvey/) and all the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey). **`SimSurvey`** can be installed via GitHub using the **`remotes`** package:

@@

TODO:
> increase the margins on the table in the template
> re-read the initial SimSurvey section and expand as necessary to better allign with the table
> in each section, expand upon the details of the core functions and the associated equation symbols (i.e. break down case-study table into several tables)
> in the case study section, copy and paste (and clean-up) the code used to generate the case study results

@@

Abundance is simulated using the `sim_abundance` function and the settings used for our case study are shown in Table \@ref(tab:par). These settings are also the defaults of the function. This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`). The package includes the functions `sim_R`, `sim_Z` and `sim_vonB` for simulating recruitment, total mortality and growth, respectively. These functions are "closures" [i.e. functions that contain data and return functions; @wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years. The use of closures also allows users to use their own closures with a similar structure but different underlying formula. Overall, the function provides a simple tool for simulating a range dynamic age-structured populations. For instance, below we simulate a relatively long (default case study settings) and short lived species.

@@

@@

The `sim_distribution` function distributes the population simulated using `sim_abundance` throughout the grid generated using `make_grid`. Table \@ref(tab:par) outlines the default settings used for the function and the case study. In addition to supplying objects produced by `sim_abundance` and `make_grid`, this function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures. Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider range) and exhibits stronger correlation across ages than years. Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.

@@

The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`. This function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey. Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively. Multiple simulations of the same survey can be run using the `n_sims` argument, however, it is important not to request large numbers of simulations because this function can quickly deplete a computers RAM. Defaults used in the function and for the case study roughly correspond to the assumed catchability in NAFO division 3Ps as well as the sampling protocol used in the survey of this stock (Table \@ref(tab:par)). Below we use `sim_survey` to simulate two surveys over our case study population, of which one is set-up to have higher sampling effort than the other.

@@

# Appendix 1: case study

The construction of the **`SimSurvey`** package was motivated by a regional need to assess the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada. As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region. Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package. Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey. Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use. See Table \@ref(tab:par) for a description of the code used to generate the case study results.

Table: (\#tab:par) Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here. Note that the settings below largely correspond with the defaults used in the functions.

|                                       **Function call** |                                            **Description** |                 **Symbol** |
|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
| `>`                                                     |                                                            |                            |
| `> abundance <- sim_abundance(`                         |                                       *Simulate abundance* |                            |
| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
| `+                     log_sd = 0.1,`                   |     Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
| `>`                                                     |                                                            |                            |
| `> grid <- make_grid(`                                  |                                              *Make a grid* |                            |
| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |
| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
| `>`                                                     |                                                            |                            |
| `> distribution <- sim_distribution(`                   |                                    *Simulate distribution* |                            |     
| `+   sim = abundance,`                                  |                Object produced by `sim_abundance` function |                            |     
| `+   grid = grid,`                                      |                    Object produced by `make_grid` function |                            |
| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
| `>`                                                     |                                                            |                            |
| `> surveys <- expand_surveys(`                          |                       *Define a series of surveys to test* |                            |
| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
| `>`                                                     |                                                            |                            |
| `> tests <- test_surveys(`                              |                             *Simulate and analyze surveys* |                            |
| `+   sim = distribution,`                               |             Object produced by `sim_distribution` function |                            |
| `+   surveys = surveys,`                                |               Object produced by `expand_surveys` function |                            |
| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |

@@

# Appendix 2: age-year-space covariance

>>>>

The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)). The equations behind these functions are detailed in the sections below. **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. Package documentation has been published online using **`pkgdown`** (https://paulregular.github.io/SimSurvey/) and all the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey). **`SimSurvey`** can be installed via GitHub using the **`remotes`** package:

@@

TODO:
> in each section, expand upon the details of the core functions and the associated equation symbols (i.e. break down case-study table into several tables)
> in the case study section, copy and paste (and clean-up) the code used to generate the case study results

@@

Table: (\#tab:abundance) Default `sim_abundance` function call, with argument descriptions and associated parameter symbols.

| **Function call**                           | **Description**                                            | **Symbol**                 |
|:------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                     |
| `+  ages = 1:20,`                           | Ages                                                       | $a$                        |
| `+  years = 1:20,`                          | Years                                                      | $y$                        |
| `+  R = sim_R(mean = 30000000,`             | Mean recruitment                                           | $\mu_r$                    |
| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                 |
| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                    |
| `+            log_sd = 0.2,`                | Standard deviation of total mortality                      | $\sigma_Z$                 |
| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, age}$    |
| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, year}$   |
| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$               |
| `+                    L0 = 5,`              | Length at birth (cm)                                       | $l_0$                      |
| `+                    K = 0.1,`             | Growth rate                                                | $K$                        |
| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                 |
| `+                    length_group = 3))`   | Length group for abundance at length (cm)                  | $l_{group}^{N}$            |

Abundance is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance). This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`). The package includes the functions `sim_R`, `sim_Z` and `sim_vonB` for simulating recruitment, total mortality and growth, respectively. These functions are "closures" [i.e. functions that contain data and return functions; @wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years. The use of closures also allows users to use their own closures with a similar structure but different underlying formula. Overall, the function provides a simple tool for simulating a range dynamic age-structured populations. For instance, below we simulate a relatively long (default case study settings) and short lived species.

@@

Table: (\#tab:distribution) Default `sim_distribution` function call, with argument descriptions and associated parameter symbols.

| **Function call**                                       | **Description**                                       | **Symbol**            |
|:------------------------------------------------------- |:----------------------------------------------------- | :-------------------- |
| `sim_distribution(`                                     | &nbsp;                                                | &nbsp;                |
| `+  sim,`                                               | Simulated population from `sim_abundance`             | &nbsp;                |
| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                      | &nbsp;                |
| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                      | &nbsp;                |
| `+                   res = c(3.5, 3.5),`                | Grid resolution (km)                                  | $R$                   |
| `+                   shelf_depth = 200,`                | Shelf depth (km)                                      | $d_{shelf}$           |
| `+                   shelf_width = 100,`                | Shelf width (km)                                      | $w_{shelf}$           |
| `+                   depth_range = c(0, 1000),`         | Depth range (m)                                       | $[d_{min}, d_{max}]$  |
| `+                   n_div = 1,`                        | Number of divisions                                   | $N_{div}$             |
| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata            | &nbsp;                |
| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution     | $\sigma_{\xi}$        |
| `+                            range = 300,`             | Range of spatial correlation (km)                     | $r$                   |
| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution       | $\varphi_{\xi, age}$  |
| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution      | $\varphi_{\xi, year}$ |
| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages  | &nbsp;                |
| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)     | $\mu_d$               |
| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)         | $\sigma_d$            |

@@

The `sim_distribution` function distributes the population simulated using `sim_abundance` throughout the grid generated using `make_grid`. Table \@ref(tab:distribution) outlines the default settings used for the function. In addition to supplying objects produced by `sim_abundance` and `make_grid`, this function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures. Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider range) and exhibits stronger correlation across ages than years. Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.

@@

Table: (\#tab:survey) Default `sim_survey` function call, with descriptions and associated parameter symbols of key arguments.

| **Function call**                   | **Description**                                                         | **Symbol**     |
|:----------------------------------- |:----------------------------------------------------------------------- | :------------- |
| `sim_survey(`                       | &nbsp;                                                                  | &nbsp;         |
| `+  sim,`                           | Simulated spatial population from `sim_distribution`                    | &nbsp;         |
| `+  n_sims = 1`                     | Number of times to repeat the survey                                    | &nbsp;         |
| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                             | $k$            |
| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$          |
| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{trawl}$    |
| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;         |
| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{sets}$     |
| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{lengths}$  |
| `+  length_group = 1,`              | Length group for age sampling (cm)                                      | $l_{group}$    |
| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{ages}$     |
| `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;         |

The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`. A default function call is described in Table \@ref(tab:survey)). This function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey. Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively. Multiple simulations of the same survey can be run using the `n_sims` argument, however, it is important not to request large numbers of simulations because this function can quickly deplete a computers RAM. Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher sampling effort than the other.

@@

# Appendix 1: Case study

The construction of the **`SimSurvey`** package was motivated by a regional need to assess the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada. As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region. Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package. Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey. Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use. The code below will replicate the results of this caset study:

set.seed(438)

abundance <- sim_abundance(
  ages = 1:20,
  years = 1:20,
  R = sim_R(mean = 30000000,
            log_sd = 0.5),
  Z = sim_Z(mean = 0.5,
            log_sd = 0.2,
            phi_age = 0.9,
            phi_year = 0.5),
  growth = sim_vonB(Linf = 120, 
                    L0 = 5, 
                    K = 0.1,
                    log_sd = 0.1, 
                    length_group = 3)
)

distribution <- sim_distribution(
  abundance,
  grid = make_grid(x_range = c(-140, 140),
                   y_range = c(-140, 140),
                   res = c(3.5, 3.5),
                   shelf_depth = 200,
                   shelf_width = 100,
                   depth_range = c(0, 1000),
                   n_div = 1,
                   strat_breaks = seq(0, 1000, by = 40),
                   strat_splits = 2),
  ays_covar = sim_ays_covar(sd = 2.8,
                            range = 300,
                            phi_age = 0.5,
                            phi_year = 0.9,
                            group_ages = 5:20),
  depth_par = sim_parabola(mu = 200,
                           sigma = 70)
)

surveys <- test_surveys(
  distribution,
  surveys = expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
                           lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
                           ages_cap = c(2, 5, 10, 20, 50)),
  n_loops = 1000,
  q = sim_logistic(k = 2, x0 = 3)
)

@@

# Appendix 2: Age-year-space covariance
